---
title: "ChatGPT Analysis of Mean Discharge in a River"
author: "Matt Ross"
date: "2023-01-23"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r R Setup}
library(tidyverse)
library(httr)
library(jsonlite)
```

# ChatGPT, the analyst

For this assignment you will be asking ChatGPT to conduct an analysis of
mean river discharge for any river station in the USA. You can use the
Poudre River in Fort Collins if you want, but I encourage you to find
other USGS stations that you are interested in. This assignment will
have three prompts for ChatGPT, and you will need to troubleshoot this
code for each prompt. The prompts are:

1.  *Using R, Download daily discharge data for XX river in YY city from
    1990-2022 (you pick the river and the city).* You can choose to
    provide ChatGPT with an actual USGS station ID which will produce
    less errors, or you can see how well it figures out station ids (its
    bad at this for some reason). You may also need to provide it with
    specific package recommendations (as I do in the videos).

2.  How would I take the summertime (June 21 - Sept 21) annual mean of
    this data and plot it with year on the x-axis and summertime mean
    flow on the y axis.

3.  How would I Use a linear model to determine if summer annual mean
    discharge is changing?

4 (Bonus). Why might a linear model be a poor choice for this analysis?

## Data Download

```{r}
library(dataRetrieval)

# Define the start and end dates for the data query
start <- "1990-01-01"
end <- "2022-12-31"

# Retrieve the data for site 01372043
HudsonData <- readNWISdv(siteNumber = "01358000",
                         parameterCd = "00060",
                         startDate = start,
                         endDate = end)

# Print the data
print(HudsonData)

```

## Data analysis and visualization

"You can use R to take the summertime annual mean of the data and plot
it with year on the x-axis and summertime mean flow on the y axis by
following these steps:

Subset the data to include only the dates between June 21 and September
21"

```{r}
# Convert the date column to a Date class
HudsonData$datetime <- as.Date(HudsonData$Date, format = "%Y-%m-%d")


# Extract data between June 21 and September 21 for all years
summer_data <- HudsonData[HudsonData$Date >= as.Date("1990-06-21") & HudsonData$Date <= as.Date("2022-09-21") & year(HudsonData$Date) >= 1990 & year(HudsonData$Date) <= 2022, ]

```

Group the data by year and calculate the annual mean flow for the
summertime

```{r}
# Group the data by year
summer_data_by_year <- summer_data %>% 
  mutate(year = year(datetime), 
         months = month(Date)) %>%
  filter(months %in% 6:10) %>%
  group_by(year) %>%
  summarise(mean_flow = mean(X_00060_00003,
                             na.rm = T))

```

Plot the data using the "ggplot2" package

```{r}
library(ggplot2)

# Plot the data
ggplot(summer_data_by_year, aes(x =year, y = mean_flow)) + 
  geom_line() +
  xlab("Year") +
  ylab("Summertime Mean Flow") +
  ggtitle("Summertime Annual Mean Flow")

```

## Model output

You can use a linear model in R to determine if the summer annual mean
discharge is changing by following these steps:

Group the data by year and calculate the annual mean flow for the
summertime

```{r}

# Calculate the annual mean flow (we did this earlier, so I've #'d it)
# annual_mean_flow <- summer_data_by_year %>% 
# summarise(mean_flow = mean(flow))

```

Fit a linear model to the annual mean flow data

```{r}
# Fit a linear model
model <- lm(data = summer_data_by_year, mean_flow ~ year)

# Summarize the model and find the p-value
summary(model)
```

P-value = 0.9752

Plot the results

```{r}
# Plot the results
library(ggplot2)

ggplot(summer_data_by_year, aes(x = year, y = mean_flow)) +
  geom_point() + 
  geom_line(aes(y = predict(model))) +
  xlab("Year") +
  ylab("Summertime Mean Flow") +
  ggtitle("Summertime Annual Mean Flow")
```

## Bonus - why linear models are dangerous time-series analysis tools.

When asked "why might a linear model be a bad choice for analysis for
this dataset?" Chat GPT responds with this:

*"A linear model may be a bad choice for analysis of this dataset for
several reasons:*

*Non-linear trends: If the data has a non-linear trend over time, a
linear model may not be able to capture the underlying patterns in the
data. A more appropriate model in this case would be a non-linear
regression model or a more complex time series model.*

*Autocorrelation: If the summertime annual mean discharge is
autocorrelated, meaning that the value at a given time is related to the
value at a previous time, then a linear model may not be appropriate. In
this case, a time series model such as ARIMA or SARIMA would be more
appropriate.*

*Lack of independence: If the summertime annual mean discharge values
are not independent, meaning that they are related to each other, then a
linear model may not be appropriate. In this case, a time series model
that accounts for the dependence structure of the data would be more
appropriate.*

*Outliers or extreme values: If the data contains outliers or extreme
values, a linear model may not be appropriate, as it assumes that the
data follows a normal distribution and may be influenced by these
outliers or extreme values. In this case, a more robust regression
method, such as quantile regression, would be more appropriate.*

*These are some of the reasons why a linear model may be a bad choice
for analysis for this dataset. However, the best choice of model will
depend on the specifics of the data and the research question being
asked, so it's important to carefully consider the underlying patterns
in the data before selecting a model."*

Chat GPT is correct on a few points with this dataset: not only is the data not independent and autocorrelated (ie: the value of a data point is related to the values preceding and following it chronologically), but it also contains an outlier that can be seen in the graphs in the late 90s, making a linear model a poor choice of analysis for this data.
